\Large{Adversarial Training against Systematic Uncertainty}

### 2016-09-29 tex/paper.tex
save

\begin{figure*}
    \begin{center}
        \includegraphics[width=0.48\textwidth]{figures/f-plain.pdf}
        \includegraphics[width=0.48\textwidth]{figures/surface-plain.pdf}\\
        \vspace{-0.8cm}
        \includegraphics[width=0.48\textwidth]{figures/f-adversary.pdf}
        \includegraphics[width=0.48\textwidth]{figures/surface-adversary.pdf}
    \end{center}
    \vspace{-1cm}
    \caption{Toy example.
    (Upper left) Conditional probability densities of the decision scores for $Z=0, 1, 2$
       when $f$ is built without adversarial training. The resulting densities
       are clearly dependent on $Z$, indicating that $f$ is not pivotal.
    (Upper right) The corresponding decision surface, highlighting
       the fact that samples are easier to classify for values of $Z$ closer to $2$,
       hence partially explaining the dependency.
    (Lower left) Conditional probability densities of the decision scores for $Z=0, 1, 2$ when $f$ is
       built with adversarial training, as outlined in Section~\ref{sec:method}.
       The resulting densities are now much closer to each other, indicating only a
       small dependency on $Z$.
    (Lower right) The corresponding decision surface, illustrating how adversarial
       training bends the decision function to erase the dependency on $Z$.
    }
    \label{fig:toy}
\end{figure*}

\begin{figure}
    \includegraphics[width=0.48\textwidth]{figures/training.pdf}
    \vspace{-1cm}
    \caption{Toy example. Training curves for ${\cal L}_f(\theta_f)$, ${\cal L}_r(\theta_f, \theta_r)$
             and ${\cal L}_f(\theta_f) - \lambda {\cal L}_r(\theta_f, \theta_r)$.
             Adversarial training was performed for $300$ iterations, mini-batches of size $M=128$, $K=10$ and $\lambda=10$.}
    \label{fig:toy-training}
\end{figure}

@@

By contrast, when running Algorithm~\ref{alg:adversarial-training} as initialized
with the classifier obtained previously, and in combination with a mixture density
network $r$ (modeling $Z|f(X;\theta_f)$ as a truncated mixture of four gaussian
components), adversarial training effectively reshapes the decision function
$f$ so it that becomes almost independent on $Z$, as shown in the lower plots
of Figure~\ref{fig:toy}. In particular, the conditional probability densities of
the decision scores $f(X;\theta_f)$ are now very similar to each other,
indicating only a small  dependency on the nuisance, as theoretically expected.
The dynamics of adversarial training is illustrated in
Figure~\ref{fig:toy-training}, where the losses ${\cal L}_f$, ${\cal L}_r$ and
${\cal L}_f - \lambda {\cal L}_r$ are evaluated after each iteration of
Algorithm~\ref{alg:adversarial-training}. In the first iterations, we observe
that the global objective ${\cal L}_f - \lambda {\cal L}_r$ is minimized by
making the classifier less optimal, hence the corresponding increase of ${\cal
L}_f$, but which results in a classifier that is more pivotal, hence the
corresponding increase of ${\cal L}_r$ and the total net benefit. As learning
goes, minimizing $E$ requires making predictions that are more accurate, hence
decreasing ${\cal L}_f$, or that are less dependent on $Z$, hence shaping
$p_{\theta_r}$ so that it becomes more uniform. Indeed, ${\cal L}_f$ eventually
start to slighlty decrease, while remaining lower bounded by $\min_{\theta_f}
{\cal L}_f(\theta_f)$ as approximated by the dashed line in the first plot.
Similarly,  ${\cal L}_r$ keeps tending towards the differential entropy $H(p_Z) =
\log(2)$ of the uniform distribution over ${\cal Z}$, as shown by the dashed
line in the second plot. Finally, let us note that the ideal situation of a
classifier that is both optimal and pivotal appears to be unreachable for this
problem, as shown in third plot by the  offset between ${\cal L}_f - \lambda
{\cal L}_r$ and the dashed line approximating $\mathbb{E}_{X} [ H(p_{Y|X}) -
\lambda H(p_Z)]$.

>>>>

\begin{figure*}
    \begin{center}
        \includegraphics[width=0.48\textwidth]{figures/f-plain.pdf}
        \includegraphics[width=0.48\textwidth]{figures/surface-plain.pdf}\\
        \vspace{-0.8cm}
        \includegraphics[width=0.48\textwidth]{figures/f-adversary.pdf}
        \includegraphics[width=0.48\textwidth]{figures/surface-adversary.pdf}
    \end{center}
    \vspace{-1cm}
    \caption{Toy example.
    (Upper left) Conditional probability densities of the decision scores for $Z=0, 1, 2$
       when $f$ is built without adversarial training. The resulting densities
       are clearly dependent on $Z$, indicating that $f$ is not pivotal.
    (Upper right) The corresponding decision surface, highlighting
       the fact that samples are easier to classify for values of $Z$ closer to $2$,
       hence partially explaining the dependency.
    (Lower left) Conditional probability densities of the decision scores for $Z=0, 1, 2$ when $f$ is
       built with adversarial training, as outlined in Section~\ref{sec:method}.
       The resulting densities are now almost identical to each other, indicating only a
       small dependency on $Z$.
    (Lower right) The corresponding decision surface, illustrating how adversarial
       training bends the decision function to erase the dependency on $Z$.
    }
    \label{fig:toy}
\end{figure*}

\begin{figure}
    \includegraphics[width=0.48\textwidth]{figures/training.pdf}
    \vspace{-1cm}
    \caption{Toy example. Training curves for ${\cal L}_f(\theta_f)$, ${\cal L}_r(\theta_f, \theta_r)$
             and ${\cal L}_f(\theta_f) - \lambda {\cal L}_r(\theta_f, \theta_r)$.
             Adversarial training was performed for $75$ iterations, mini-batches of size $M=128$, $K=100$ and $\lambda=50$.}
    \label{fig:toy-training}
\end{figure}

@@

Let us now consider the adversarial training of $f$ against an adversary $r$
implemented as a mixture density network modeling $Z|f(X;\theta_f)$ as a
truncated mixture of four gaussians over ${\cal Z}$. When running
Algorithm~\ref{alg:adversarial-training} as initialized with the classifier $f$
obtained previously, adversarial training effectively reshapes the decision
function so it that becomes almost independent on the nuisance parameter, as
shown in the lower plots of Figure~\ref{fig:toy}. In particular, the conditional
probability densities of the decision scores $f(X;\theta_f)$ are now very
similar to each other, indicating only a small  dependency on the nuisance, as
theoretically expected. The dynamics of adversarial training is illustrated in
Figure~\ref{fig:toy-training}, where the losses ${\cal L}_f$, ${\cal L}_r$ and
${\cal L}_f - \lambda {\cal L}_r$ are evaluated after each iteration of
Algorithm~\ref{alg:adversarial-training}. In the first iterations, we observe
that the global objective ${\cal L}_f - \lambda {\cal L}_r$ is minimized by
making the classifier less accurate, hence the corresponding increase of ${\cal
L}_f$, but which results in a classifier that is more pivotal, hence the
corresponding increase of ${\cal L}_r$ and the total net benefit. As learning
goes, minimizing $E$ then requires making predictions that are more accurate,
hence decreasing ${\cal L}_f$, or that are less dependent on $Z$, hence shaping
$p_{\theta_r}$ so that it becomes more uniform. Indeed, ${\cal L}_f$ eventually
start to slightly decrease, while remaining lower bounded by $\min_{\theta_f}
{\cal L}_f(\theta_f)$ as approximated by the dashed line in the first plot.
Similarly,  ${\cal L}_r$ keeps tending towards the differential entropy $H(p_Z) =
\log(2)$ of the uniform distribution over ${\cal Z}$, as shown by the dashed
line in the second plot. Finally, let us note that the ideal situation of a
classifier that is both optimal and pivotal appears to be unreachable for this
problem, as shown in third plot by the  offset between ${\cal L}_f - \lambda
{\cal L}_r$ and the dashed line approximating $\mathbb{E}_{X} [ H(p_{Y|X}) -
\lambda H(p_Z)]$.
